### 2025-12-10 submissions/Earlier Versions of Manuscript/manuscript2.tex

\documentclass[%
 reprint,
superscriptaddress,
groupedaddress,
%unsortedaddress,
%runinaddress,
%frontmatterverbose, 
%preprint,
%preprintnumbers,
%nofootinbib,
%nobibnotes,
%bibnotes,
 amsmath,amssymb,
 aps,
prl,superscriptaddress
%prb,
%rmp,
%prstab,
%prstper,
%floatfix,
]{revtex4-2}

\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{braket}
\usepackage{color}
\graphicspath{{./figures/}}
\bibliographystyle{apsrev4-2}

\begin{document}

%\title{Pseudogapped non-Fermi liquid phase arising from Kondo breakdown at the Mott transition}

%\title{Long range entangled Fermi arcs at a Pseudogapped Mott transition}

\title{Pseudogap and Mott criticality: Stretching Kondo Screening to Breaking Point}

%\title{What lies between a Fermi liquid and a Mott insulator in two dimensions? Insights from an impurity model}% Force line breaks with \\

\author{Abhirup Mukherjee}
\affiliation{%
 Department of Physical Sciences, Indian Institute of Science Education and Research Kolkata, Nadia - 741246, India
}
\author{S. R. Hassan}
\affiliation{The Institute of Mathematical Sciences, C.I.T. Campus, Chennai 600 113, India}
\author{Anamitra Mukherjee}
\affiliation{School of Physical Sciences, National Institute of Science, Education and Research, HBNI, Jatni 752050, India}
\author{N. S. Vidhyadhiraja}
\affiliation{Theoretical Sciences Unit, Jawaharlal Nehru Center for Advanced Scientific Research, Jakkur, Bengaluru 560064, India}
\author{A. Taraphder}
\affiliation{Department of Physics, Indian Institute of Technology Kharagpur, Kharagpur 721302, India}
%
\author{Siddhartha Lal}
\affiliation{%
 Department of Physical Sciences, Indian Institute of Science Education and Research Kolkata, Nadia - 741246, India
}

\date{\today}
\begin{abstract}
We present an auxiliary model approach that illustrates the centrality of the pseudogap phenomenon in the Mott transition of the Hubbard-Heisenberg model on a half-filled square lattice. The passage from a Fermi liquid to a Mott insulator involves two stages: the first involves a systematic unravelling of Kondo screening, frustrated by charge fluctuations in the conduction bath of the underlying lattice-embedded quantum impurity model. This destabilises the Fermi liquid, culminating in the disconnection of the Fermi surface with the appearance of gaps at the antinodes. In the second stage, an emergent pseudogap phase arises from an effective two-channel Kondo problem in the underlying impurity model. The resulting Fermi arcs possess non-Fermi liquid excitations proximate to a critical Fermi surface. Upon approaching the Mott transition, the nature of this exotic metal evolves with shrinking of the Fermi arcs towards a singular nodal non-Fermi liquid. This accompanies the onset of long-range real-space entanglement and spin-flip correlations, revealing the Mott transition on the square lattice to be beyond the paradigm of local quantum criticality.
\end{abstract}

\maketitle

\par\noindent\textit{Introduction.}
The origin of the pseudogap and strange metal phases of the cuprates continues to be hotly debated in the context of high-temperature superconductivity~\cite{keimer2015quantum}. The anomalous pseudogap (PG) phase, observed in the cuprates, exhibits nodal-antinodal dichotomy with spectral weight concentrated on Fermi arcs around the nodal regions~\cite{loeserKapitulnik1996,Norman1998,Hashimoto2014}. While several calculations have observed a PG in the two dimensional Hubbard model ~\cite{KyungKotliar2006,MacridinAzevedo2006,WuFerrero2018,anirbanmott2,HilleAndergassen2020}, there is no general consensus regarding various aspects of the PG phase, including its relation to the proximate Mott insulating and superconducting phases~\cite{FradkinRevModPhys2015,anirbanmott2,Kitatani2023,Sorella2023}, its evolution from weak- to strong-coupling ~\cite{HuangDevereaux2018,Fedor2022,SimkovicFerrero2024}, and whether the nodal-antinodal dichotomy is intrinsic to the model~\cite{Hashimoto2014,Schafer2021}. Importantly, the connection between finite-temperature crossovers and zero-temperature ground states observed in various analyses remains unclear~\cite{White1998,Ido2018,ProustTaillefer2019,Ponsioen2019,Shengtao2021,XuZhang2022}.

\textcolor{blue}{Viewed as a quantum mechanical phenomenon, the PG involves the emergence of disconnected gapped regions of the Fermi surface (known as Luttinger surfaces of zeros of the single-particle Greens function (CITE references)) from a Fermi liquid with a connected Fermi surface. The growth and eventual connection of these Luttinger surfaces then leads to the transition into the Mott insulator. Which physical principles guide these events and what do they teach us about the criticality that obtains Mott insulation? Further, can the excitations of the gapless Fermi arcs be adiabatically connected to those of a non-interacting electron gas in the presence of the Luttinger surfaces? If not, what kind of scale-invariant quantum matter do they characterise? Which features of the PG elucidate the breakdown of screening in the Fermi liquid and lead instead to a singular self-energy in the Mott insulator?}

We present a new auxiliary model approach 
to the strong coupling physics of particle-hole symmetric 2D Hubbard-Heisenberg model at zero temperature, revealing the origin and nature of its PG phase and the subsequent Mott transition. The first step 
requires identifying an appropriate lattice-embedded quantum impurity model, and a scaling analysis of its low-energy phases. Subsequently, a lattice model is  constructed by applying many-body lattice translation operators~\cite{stoyanova} on the impurity model (henceforth referred as \textit{tiling}), obtaining thereby various quantities of the former from the latter. Translation invariance is thus restored differently within our approach from the dynamical mean-field approximation~\cite{georges1996} and its cluster variants~\cite{Hettler2000DCA,Kotliar2001Cellular,Maier2005,Sakai2023}.

Combining the scaling analysis of the impurity model with tiling procedure provides momentum-space resolution crucial to unveiling the PG phenomenon, and clarifies the role of the Kondo breakdown process at the heart of Mott transition~\cite{fabrizio2017kondophysicsmotttransition,RADEMAKER2025}. 
Emerging from a systematic unraveling of Kondo screening, the PG phase is separated from the Fermi liquid (FL) and Mott insulating phases by correlation-driven Lifshitz transitions of the Fermi surface (FS)~\cite{sakai2009evolution}. Passage through the PG involves dynamical transfer of spectral weight between gapped anti-nodal Luttinger surfaces (zeros of the single-particle Green's function) and gapless nodal arcs with non-FL excitations. We also observe the \textcolor{blue}{concomitant} onset of long-range real-space entanglement and spin-flip correlations close to the Mott transition, rendering it beyond local quantum criticality~\cite{Si2001}. \textcolor{blue}{Our findings serve as a demonstration of a recent proposal (CITE La Nave etal. arXiv:2506.04342) that disconnected Luttinger surfaces leads to the spontaneous violation of a $\mathbb{Z}_{2}$ symmetry associated with the Fermi surface (CITE HALDANE-ANDERSON, Recent work by Philips etal), and thence to the existence of non-local propagating degrees of freedom.}

\par\noindent\textit{A lattice-embedded impurity model.}
Building on the insight, that the Kondo breakdown in a simple extension of the Anderson impurity model~\cite{Mukherjee_2023} captures the physics of Mott transition in infinite dimensions~\cite{georges1996}), we study a similar Anderson impurity model embedded within a square lattice 
\begin{equation}\begin{aligned}\label{impurityModel}
	\mathcal{H}_\text{aux} = H_\text{imp} + H_\text{imp-cbath} + H_\text{cbath}~, 
\end{aligned}\end{equation}
where $H_{imp} = - \frac{U}{2}\left(\hat n_{d \uparrow} - \hat n_{d \downarrow} \right) ^2$ is the Hamiltonian for the localised impurity site, and the impurity (with creation operator $c^\dagger_{d\sigma}$, and spin-1/2 moment ${\bf S}_d$) interacts with four nearest neighbour sites ($c_{Z\sigma}$, labelled by index $Z$) 
through local Kondo terms of uniform strength $J$ as well as single-particle hybridisation $V$, \(H_\text{imp-cbath} =\frac{1}{2} J\sum_{\sigma_1,\sigma_2}\sum_{Z} {\bf S}_d\cdot c^\dagger_{Z\sigma_1}{\boldsymbol \tau}_{\sigma_1,\sigma_2} c_{Z\sigma_2} - V \sum_{\sigma, Z} \left(c^\dagger_{Z\sigma} c_{d\sigma} + h.c.\right)\). Finally, the term
$H_\text{cbath} = \sum_{{\bf k}}\epsilon_{\textbf{k}}
c^\dagger_{{\bf k},\sigma}c_{{\bf k},\sigma} -\frac{W}{2}\sum_{Z} \left(n_{Z\uparrow} - n_{Z\downarrow}\right)^2$ includes the kinetic energy ($\epsilon_{\textbf{k}}=-2t\left(\cos k_x + \cos k_y\right)$) for tight-binding conduction bath electrons ($c^\dagger_{{\bf k},\sigma}$) on a half-filled square lattice, together with local correlations ($W$) that frustrate Kondo screening. The 2D structure of the impurity model and its distinction from the infinite-dimensional counterpart becomes apparent on Fourier transforming the Kondo coupling to ${\bf k}$-space:
$J_{{\bf k}, {\bf k}^\prime} = \frac{J}{2}\left[\cos\left({\bf k}_x - {\bf k}^\prime_x\right) + \cos\left({\bf k}_y - {\bf k}^\prime_y\right)\right]$, with $J>0$. Clearly, $J_{{\bf k}, {\bf k}^\prime}$ possesses $C_{4}$ lattice symmetry. 
\par\noindent\textit{Reconstructing the lattice model.}
We now define the {\it tiling} procedure, which recreates the lattice model from the impurity Hamiltonian. For a {\it unit cell} of the tiling,  
we consider \(\mathcal{H}_\text{aux}({\bf r}_d)\) to have the impurity at a reference site \({\bf r}_d\) of our lattice. To create the lattice model, we translate the unit cell across all sites of the lattice:
\begin{equation}\begin{aligned}
	\label{tilingPrescription}
	\mathcal{H}_\text{tiled} &= \sum_{{\bf r}}T^\dagger({\bf r})\mathcal{H}_\text{aux}({\bf r}_d)T({\bf r}) - N\mathcal{H}_\text{cbath},
\end{aligned}\end{equation}
where \({\bf r}\) sums over all lattice sites, and $T({\bf r})$ is a many-body translation operator that translates all coordinates by a vector \({\bf r}\). Details of tiling procedure are provided in the Supplemental Information (SI)~\cite{suppmat}. Implementing this for the extended Anderson impurity model (eq.~\ref{impurityModel}) generates a {\it Hubbard-Heisenberg} lattice model in two spatial dimensions:
    \begin{eqnarray}
        &&\mathcal{H}_\text{tiled} = -\frac{\tilde t}{\sqrt\mathcal{Z}}\sum_{\left<{\bf r}_i, {\bf r}_j\right>;\sigma}\left(c^\dagger_{{\bf r}_i,\sigma}c_{{\bf r}_j,\sigma} + \text{h.c.}\right) - \tilde \mu \sum_{{\bf r}}\hat n_{{\bf r},\sigma}\nonumber \\
        &&+ \frac{\tilde J}{\mathcal{Z}}\sum_{\left< {\bf r}_i, {\bf r}_j\right>}{\bf S}_{{\bf r}_i}\cdot{\bf S}_{{\bf r}_j} - \frac{1}{2}\tilde U\sum_{\bf r}\left(\hat n_{{\bf r} , \uparrow} - \hat n_{{\bf r} , \downarrow}\right)^2~,
        \end{eqnarray}
with parameters $(\tilde{t},\tilde{\mu},\tilde{J},\tilde{U})$ that are related simply to those in $\mathcal{H}_\text{aux}$~\cite{suppmat}. The eigenstates of $H_{\text{tiled}}$ are dictated by a many-body version of Bloch's theorem~\cite{stoyanova}. Relations between Hamiltonians and eigenstates of the impurity and lattice models are exploited to obtain equal-time correlation functions and entanglement measures of the latter from the former~\cite{suppmat}.
\begin{figure}
    \centering
    \includegraphics[width=0.31\linewidth, height=2.45cm]{phaseDiagram.pdf}
    \includegraphics[width=0.67\linewidth]{SF-DOS.pdf}
    \caption{Left: Phase diagram of impurity model at strong coupling in $U$ in terms of competing dimensionless Kondo ($J/t$) and bath correlation ($W/t$) couplings. A pseudogap phase (red, PG) is observed between the local Fermi liquid (pink, LFL) and local moment (black, LM) phases. Center: $k$-space-resolved spin-spin correlation $\chi_s(d,{\bf k}) = \braket{{\bf S}_d\cdot{\bf S}_{\bf k}}$ in the pseudogap phase of the impurity model. Antinodal regions are observed to decouple from Kondo screening of the impurity. Right: Upon tiling, this leads to a 
    $k$-space-resolved antinodal gap in the electronic density of states of the lattice model, corresponding to Luttinger surfaces of zeros.}
    \label{spinCorr}
\end{figure}
%\par\noindent\textit{Pseudogapping{\textcolor{red}``pseudogapping"  DOESN'T SOUND GOOD}} transition from Kondo breakdown.}

\par\noindent\textit{Pseudogap Formation via Kondo Breakdown.}
A detailed picture of Kondo breakdown in the large $U$ phase of $H_\text{aux}$ is obtained from a scaling analysis of the ${\bf k}$-resolved Kondo coupling \(J^{(j)}_{{\bf k}_1, {\bf k}_2}\) using the 
unitary renormalisation group (URG) method \cite{anirbanurg1} (see \cite{suppmat} for details)
\begin{equation}\begin{aligned}\label{KondoRGequation}
	\Delta J^{(j)}_{{\bf k}_1, {\bf k}_2} = -\sum_{{\bf q} \in \text{PS}} \frac{J^{(j)}_{{\bf k}_2,{\bf q}} J^{(j)}_{{\bf q},{\bf k}_1} + 4J^{(j)}_{{\bf q}, {\bf \bar q}} W_{{\bf \bar q}, {\bf k}_2, {\bf k}_1, {\bf q}}}{\omega - \frac{1}{2}|\varepsilon_j| + J^{(j)}_{{\bf q}}/4 + W_{{\bf q}}/2}~,
\end{aligned}\end{equation}
where \(\varepsilon_j\) is the energy of the shell being decoupled at the \(j^\text{th}\) step, the sum is over all occupied momentum states \({\bf q}\) 
%in the particle sector (PS {\textcolor{green}{are we using the abbreviation PS anywhere else?}}) 
of the energy shell \(\varepsilon_j\), 
%(i.e., all states occupied at \(T=0\) and in the absence of any quantum fluctuations), 
and  \({\bf \bar q} = {\bf q} + {\boldsymbol \pi}\) is the particle-hole transformed state associated with ${\bf q}$. The bath interaction coupling $W_{{\bf \bar q}, {\bf k}_2, {\bf k}_1, {\bf q}}$ is found to be marginal under these transformations. While the complexity of the RG equation for $J_{{\bf k}_1, {\bf k}_2}$ (eq.\eqref{KondoRGequation}) is clarified through a detailed numerical evaluation whose results are discussed below, its structure permits the broad conclusion that the frustration of Kondo screening due to charge fluctuations (for attractive bath interactions $W<0$) lead to the Mott metal-insulator transition~\cite{Mukherjee_2023}.

Upon tuning the ratio of the bath and Kondo interactions ($W/J$) from zero to negative values (see phase diagram in Fig.\ref{spinCorr}(left)), the following phases emerge in the impurity model from the competition between $J$ and $W$ in eq.~\eqref{KondoRGequation}: (i) for $W/J<(W/J)_{\text{PG}}$, an LFL phase (where the entire FS participates in Kondo screening), (ii) for $\frac{W}{J} \in [(\frac{W}{J})_{\text{PG}}, (\frac{W}{J})_c]$, a local PG phase where disconnected parts of the FS around the node participate in Kondo screening, and (iii) a local moment phase for $\frac{W}{J} > (\frac{W}{J})_c$, where the impurity remains unscreened at low-energies. These can be visualised from spin correlations $\chi_s(d,{\bf k}) = \braket{{\bf S}_d\cdot{\bf S}_{\bf k}}$; see Fig.\ref{spinCorr}(center) for $\chi_s(d,{\bf k})$ in the PG. The values $(W/J)_{\text{PG}}$ and $(W/J)_c$ are therefore the entry into and exit from the PG phase. Mapping on to the lattice model via tiling, we observe that the $T=0$ Mott transition of the 2D Hubbard-Heisenberg model proceeds from FL to Mott insulator through an intervening PG phase having a partially gapped FS (i.e., a vanishing electronic density of states around the antinodal regions, Fig.\ref{spinCorr}(right)). We now outline the mechanism for the formation of the PG.

\begin{figure}[htpb]
    \centering
    \includegraphics[width=\linewidth]{zerosFlow.pdf}
    \caption{Left, Right and Center: Decoupling of $J_{{\bf k}, {\bf k'}}$ (positive, negative and zeros shown in yellow, purple and green respectively), with ${\bf k}$ (grey circle) corresponding to the node, antinode and a point mid-way between them on the top right arm of the FS respectively. This is seen via the appearance of zeros (green patches) of $J_{{\bf k}, {\bf k'}}$ for ${\bf k'}$ initially on the nodal regions of adjacent arms, and their subsequent progression towards the antinodes. The decoupling 
    ends with the onset of the pseudogap.
    }
    \label{rgProgression}
\end{figure}
%\newline 
\par\noindent\textit{Unraveling of Kondo screening.} 
As shown in Fig.\ref{rgProgression}, the anisotropy in the ${\bf k}$-space of Kondo breakdown can be visualized in terms of zeros of $J_{{\bf k}_N, {\bf k}}$, involving spin-flip scattering between the node ${\bf k}_N = (\pi/2, \pi/2)$  and a general wavevector ${\bf k}$. 
For any $W/J$, the $\mathcal{C}_{4}$ lattice symmetry dictates that $J_{{\bf k}_N, {\bf k}}$ vanishes if ${\bf k}$ belongs to any of the antinodes or adjacent nodes. Tuning $W/J$ towards $(W/J)_{\text{PG}}$ leads to an unraveling of the Kondo screening: $J_{{\bf k}_N, {\bf k}}$ for ${\bf k}$ close to the adjacent nodes turns RG-irrelevant first, and a patch of zeros subsequently appears in $J_{{\bf k}_N, {\bf k}}$ around this point (Fig.\ref{rgProgression}(left)). Tuning $W/J$ further extends the patch of zeros towards the antinodes (Fig.\ref{rgProgression}(center) and (right)). Kondo screening thus unravels by a systematic decoupling of all $J_{{\bf k}_1, {\bf k}_2}$ that connect adjacent quadrants of the Brillouin zone.   
%{{\textcolor{cyan}{Tuning $W/J$ further extends the patch of zeros in $J_{{\bf k}_1, {\bf k}_2}$ for ${\bf k}_{1}$ lying between a given node and the nearest antinodes (Fig.\ref{rgProgression}(center) and (right)), visualising a systematic decoupling of $J_{{\bf k}_1, {\bf k}_2}$ connecting adjacent quadrants of the Brillouin zone quadrants -- (CHECK THIS SENTENCE, TOO LONG AND SOMEWHAT UNCLEAR)}}.

Precisely at $W/J=(W/J)_{\text{PG}}$, the antinode joins this connected region of zeros in $J_{{\bf k}_1, {\bf k}_2}$, marking the decoupling of the antinodes from all other points in the neighbourhood of the FS. This is an interaction-driven Lifshitz transition of the FS, and marks the entry into a PG phase possessing Fermi arcs~\cite{WuFerrero2018}. Importantly, it coincides with an emergent two-channel Kondo (2CK) impurity model, where each channel corresponds to a pair of Fermi arcs on opposite faces of the conduction bath FS. The 2CK nature of the PG is guaranteed by the symmetry: 
$J_{{\bf k},{\bf k'}}= -J_{{\bf k}+{\bf Q},{\bf k'}}=-J_{{\bf k},{\bf k'}+{\bf Q}}$, where ${\bf Q} = (\pi,\pi)$. The PG expands by shrinking these disconnected Fermi arcs towards the respective nodes, leading to nodal metals whose disappearance heralds the Mott transition.

\par\noindent\textit{{\bf k}-resolved dynamical spectral weight transfer.}
 Strong charge fluctuations develop between the nodal and antinodal regions of the FS in the PG regime (Fig.\ref{chargeCorr}(left)), removing low-energy spectral weight from the antinodes to higher energies and leading to selective gap formation. Accordingly, the PG coincides with the appearance of poles of the lattice model self-energy $\Sigma ({\bf k},\omega=0)$ near the antinodes; these poles approach the nodes on tuning towards the Mott transition (Fig.\ref{chargeCorr}(right)). This mirrors the coalescing of self-energy poles at zero frequency in the underlying impurity model~\cite{suppmat}. \textcolor{blue}{This process corresponds to the formation of Luttinger surfaces of zeros of the single-particle Greens function (CITE Dzyaloshinskii PRB  68, 085113 (2003)) whose growth characterises the PG, and connection to the Mott insulator.} 
\begin{figure}
    \centering
    \includegraphics[width=0.49\linewidth]{cfnode-2.pdf}
    \includegraphics[width=0.49\linewidth]{selfEnergyKspace-3.pdf}
    \caption{Left: Enhanced charge isospin correlations $\chi_c({\bf k}_1, {\bf k}_2) = \braket{c^\dagger_{{\bf k}_1\uparrow}c^\dagger_{{\bf k}_1\downarrow}c_{{\bf k}_2\downarrow}c_{{\bf k}_2\uparrow} + \text{h.c.}}$ between the nodal and antinodal regions, signalling Kondo breakdown in the pseudogap phase of the impurity model. Right: In turn, the breakdown leads to the gapping of the antinodal regions in the lattice model, seen from the appearance of poles in the imaginary part of the self-energy.}
    \label{chargeCorr}
\end{figure}

\par\noindent\textit{Non-FL excitations within the pseudogap.}
In the PG regime, 
%$|W/J|_{\text{PG}} < |W/J| < |W/J|_{c}$, 
the nature of gapless Fermi arcs changes dramatically. We have already argued that the low-energy dynamics of these gapless Fermi arcs are governed by an underlying two-channel Kondo (2CK) impurity model~\cite{Tsvelick_weigmann_mchannel_1985,emery_kivelson}.
\textcolor{red}{The excitations of the 2CK exhibit a separation of spin, charge and channel quantum numbers~\cite{affleck1992}, with gapless excitations carrying only the spin quantum number. Previous works show that the impurity self-energy of such a two-channel Kondo model displays marginal Fermi liquid (MFL) behaviour~\cite{Coleman_tsvelik,Schofield1997,Patra2023MCK} with real and imaginary parts of the self-energy as $\Sigma^{\prime} \sim \omega\ln\omega, \Sigma^{\prime\prime}\sim-\omega$. This gives rise~\cite{varma2002singular} to a linear scattering rate \(\Gamma \sim \omega\), and a logarithmically vanishing quasiparticle residue $Z_\text{imp}(\omega) \sim (c - \ln \omega)^{-1}$ (where \(c\) is a constant).} 
This is consistent with the rapid fall of the impurity quasiparticle residue $Z_\text{imp}$ (Fig.\ref{channelDecoupling}(left)) from finite values in the FL phase to vanishingly small values in the PG. The inset of Fig.\ref{channelDecoupling}(left) shows the emergence of increasingly uncompensated local magnetic moments upon traversing the PG phase. \textcolor{blue}{The underlying impurity spectral function of the gapless arcs show a pseudogapped behaviour for $\omega\simeq 0$, with a rapid fall in the spectral weight at $\omega=0$ upon traversing the PG. The collapse of the Kondo resonance into a pseudogapped spectral function is another display of the dynamical spectral weight transfer (CITE Dzyaloshinskii PRB  68, 085113 (2003)). Concomitant with this is the emergence of a zero-frequency peak in the imaginary part of the self-energy of the NFL in the PG phase, with $\Sigma''(\omega)\sim \omega^{\alpha}$. Remarkably, we find that the exponent $\alpha=-2$ characterises the NFL for the entire PG phase (CITE FIGURE); this is in stark contrast with the vanishing $\Sigma''(\omega)\sim \omega^{2}$ for the Fermi liquid). The height of the zero-frequency peak rises by almost 4 orders of magnitude from the start of the PG till its end at the Mott transition point (CITE FIGURE INSET?); this dramatic growth of the peak height coincides with the coalescing of the finite-frequency poles of the self-energy into a single pole at zero-frequency, signalling the singular nodal non-Fermi liquid present at the Mott quantum critical point.}

\begin{figure}
    \centering
    \includegraphics[width=0.49\linewidth]{localQPResidue.pdf}
    \includegraphics[width=0.49\linewidth]{I2-di_69-2000.pdf}
    \caption{Left: Suppression of quasiparticle residue as the impurity model is tuned towards the Mott transition. An initial drastic fall observed for $W/J \lesssim (W/J)_\text{PG}$, signalling the destruction of the FL with unraveling of Kondo screening. A second drastic fall is observed close to the Mott transition due to a divergent self-energy. Inset shows the growth of unscreened impurity magnetic moment in the pseudogap phase. 
    Right: Mutual information $I_{2}(d,r)$ between the impurity spin and conduction bath local spins as a function of the distance {\bf r} between them. $I_{2}(d,r)$ decays exponentially in the FL phase (blue), but show long-ranged behaviour in the non-FL phase (green, yellow). The range of the entanglement is seen to grow yet larger particularly close to the transition (green).
    }
    \label{channelDecoupling}
\end{figure}

\par\noindent\textit{Non-local nature of the pseudogap.}
In Fig.\ref{channelDecoupling}(right), both the mutual information and spin-flip correlations between the impurity spin and conduction bath sites are observed to undergo a crossover within the PG, from a short-ranged behaviour at its onset, to a long-ranged behaviour as the Mott transition approaches~\cite{suppmat}. 
%Similar long-ranged behaviour of spin-flip correlations is also observed \textcolor{blue}{in Fig.\ref{channelDecoupling}(right)} between the impurity and bath sites~\cite{suppmat}. 
This is further evidence of the breakdown of local Kondo screening, and the resulting Landau quasiparticle excitations. 
This implies that the observed Mott transition lies beyond local quantum criticality scenario~\cite{Si2001}: the non-FL Fermi arcs of the PG phase display increasingly critical behaviour close to the transition, with dynamics that cannot be described by local quantum fluctuations, and excitations that become truly long-ranged. \textcolor{blue}{Discuss that the pseudogapped Fermi arcs and gapped Luttinger surfaces (zeros of Greens function) arise from splitting of double poles on the Fermi surface into poles on either side together with zeros pinned at the FS; gapping and emergence of zeros breaks $\mathcal{Z}_{2}$ symmetry of the FS (cite Haldane-Anderson, recent works by Phillips et al.), and disconnection of the homotopy charge associated with the Luttinger count with the physical charge, i.e., the backflow contribution from the zeros at the FS must be taken into account (for the half-filled p-h symmetric system being studied here) to account for how Luttinger's volume is maintained. Further, that the existence of this Adler-Bell-Jackiw triangle anomaly related to the PG FS leads naturally (via recent works of Casini et al., La Nave et al., Altshuler et al., von Neumann, Haag etc.) to the existence of non-local propagating degrees of freedom in the gapless NFL arcs. The appearance of zeros at the FS affects the Atiyah-Singer index associated with the FS (cite Heath and Bedell) that leads to the Luttinger volume. As the ABJ triangle anomaly is captured by Atiyah-Singer index theorem (cite Jackiw et al lectures), the appearance of this anomaly at the FS topology changing Lifshitz transition is linked to changes in topological properties of the FS, and hence invariant under renormalization. Thus, the drastic change in nature of the real-space excitations (from local Landau qps to something increasingly non-local in the NFL Fermi arcs of the PG phase) is dictated by topological principles even as they are emergent under renormalisation.}

\par\noindent\textit{Evolution of the pseudogapped non-FL.} Very close to the transition, the excitations of the marginal FL are replaced by those of a Hatsugai-Kohmoto model~\cite{Baskaran1991,Hatsugai1992}. This insight is obtained from a perturbation-theoretic treatment of the RG fixed point Hamiltonian of the impurity model for $W/J\lesssim (W/J)_{\text{PG}}$, by considering the effects of a small fixed point Kondo scattering probability \(J^*\) in the backdrop of a larger bath interaction parameter \(|W|\). This yields the HK model~\cite{Baskaran1991,Hatsugai1992} as the singular part of the effective Hamiltonian arising from forward scattering processes (details in Appendix~\ref{hkmDerivation}):
\begin{equation}\begin{aligned}\label{HKModel}
	\Delta \tilde H_{{\bf q}_1 = {\bf q}_2} = \sum_{{\bf q},\sigma}\epsilon_{{\bf q}}{n}_{{\bf q},\sigma} + u\sum_{{\bf q}, \sigma}n_{{\bf q} \sigma} n_{{\bf q} \bar\sigma}~,
\end{aligned}\end{equation}
where the number operator \(n_{{\bf q} \sigma} = \phi^\dagger_{{\bf q}, \sigma} \phi_{{\bf q}, \sigma}\) pertains to emergent fermionic relative modes $\phi_{{\bf q}, \sigma}$ that are shifted by an excitation momentum \({\bf q}\) away from the nodal point \({\bf N}_1 = \left(\pi/2, \pi/2\right)\) and its 
partner \({\bf N}_1 + {\bf Q}_1 = \left(-\pi/2, -\pi/2\right)\). The kinetic energy \(\epsilon_{\bf q}\) and interaction energy \(u\sim J^{2}/W\) are dispersion and Kondo scales renormalised by conduction bath correlations~\cite{suppmat}.

Consequently, the resulting non-FL metal of the lattice model involves long-lived excitations of multiple \({\bf k}\)-states, and manifest in the form of 
a divergent one-particle self-energy at the non-interacting FS face~\cite{Phillips2020}:
	$\Sigma_{{\bf q}}(\omega) = -\frac{u^2/4}{\omega - \epsilon_{\bf q}}$, such that $\Sigma_{\epsilon_{\bf q}=0}(\omega \to 0)  \to \infty$. This zero-frequency self-energy pole 
presages the transition into a Mott insulating phase, where it marks a hard gap in the spectral function for charge excitations. This is consistent with our findings for the lattice (Fig.\ref{chargeCorr}(right)) and impurity self-energies~\cite{suppmat}. Further, for small but non-zero values of \(\omega - \epsilon_{\bf k}\), we obtain the quasiparticle residue \(Z_\text{imp} \sim \omega^2/U^2 \), i.e., $Z_{imp}$ vanishing faster compared to MFL. This explains the sharp fall in $Z_\text{imp}$ observed at the transition (Fig.\ref{channelDecoupling}(left)). The scattering rate at the FS \(\omega=\epsilon_{\bf k}\) diverges, \(\Gamma \sim U^2\delta(\omega - \epsilon_{\bf k})\), \textcolor{red}{in contrast to \(\Gamma \sim \omega\) observed in MFL}.
 Additionally, we observe that the nodal metal possesses pairing fluctuations~\cite{suppmat} that can become dominant upon doping~\cite{Phillips2020}.
\par\noindent\textit{Conclusions.} We find compelling evidence that the Mott transition involves continuous passage through a PG phase possessing an increasingly singular, non-local nodal non-FL metal, and the onset of long-ranged entanglement at criticality. The evolution of this singular PG phase with doping 
deserves further attention.
\par\noindent\textit{Acknowledgments.}
AM thanks IISER Kolkata for funding through a JRF and SRF. S Lal thanks the SERB, Govt. of India for funding through MATRICS Grant MTR/2021/000141 and Core Research Grant CRG/2021/000852.
\bibliography{tilingProject}% Produces the bibliography via BibTeX.

\appendix

\section{Appendixes}
\paragraph*{Luttinger's theorem in the presence of Luttinger surfaces.} One can also ask whether Luttinger's theorem is satisfied in our model in the presence of Luttinger surfaces. It has been shown that particle-hole symmetric systems always satisfy a generalised Luttinger's theorem~\cite{seki2017topological}, where the Fermi volume $V_L$ is represented as the difference in the number of poles and zeros, of the single-particle Greens function, that are enclosed by the FSface~\cite{seki2017topological}. This holds true even in the presence of a divergent self-energy~\cite{Phillips2013}. Since our model is always at half-filling, Luttinger's theorem is always satisfied.

The way it works out (despite the presence of Luttinger surfaces) is as follows. We need only show that the number of occupied states below the chemical potential remains unchanged across the first Lifshitz transition. Before the transition, the presence of gapless excitations ensures that the FSface is singly-occupied on average. Inside the pseudogap, the gapless $k-$states continue to contribute one pole on average to the Luttinger count, while the gapping out of certain $k-$states leads to a rearrangement of their spectrum: doubly-occupied states on the Luttinger surfaces become more favourable compared to the singly-occupied states (due to the attractive $W$-interaction). Owing to particle-hole symmetry, these states are degenerate with the zero occupancy states, so that on average, a single state is again occupied. This ensures that the number of occupied states (and hence the number of occupied poles) remains unchanged across the transition. The same argument works for the Mott insulator, where the entire FSface gets replaced by a Luttinger surface.

\paragraph*{Tiling towards a Hubbard-Heisenberg model with an embedded extended SIAM}
We give more details here regarding the construction of the lattice model from the impurity model. As mentioned in the manuscript, the lattice-embedded impurity model is of the form $H({\bf r}_d) = H_\text{imp} + H_\text{cbath} + H_\text{imp-cbath}$, where $H_\text{imp}$ is the impurity part:
\begin{equation}\begin{aligned}\label{onsiteHamiltonian}
	H_\text{imp} = -\frac{U}{2}\left(\hat n_{{\bf r}_d \uparrow} - \hat n_{{\bf r}_d \downarrow} \right)^2 - \eta\sum_\sigma \hat n_{{\bf r}_d\sigma}~,
\end{aligned}\end{equation}
\(\mathcal{H}_\text{cbath}\) is the conduction bath part:
\begin{equation}\begin{aligned}\label{bathHamiltonian}
	{H}_\text{cbath} = -\frac{1}{\sqrt\mathcal{Z}}t\sum_{\left<{\bf r}_i, {\bf r}_j\right>\neq{\bf r}_d;\sigma}\left(c^\dagger_{{\bf r}_i,\sigma}c_{{\bf r}_j,\sigma} + \text{h.c.}\right) - \\
    \frac{1}{2 \mathcal{Z}}W\sum_{{\bf z}\in\text{NN}({\bf r}_d)}\left(\hat n_{{\bf z}, \uparrow} - \hat n_{{\bf z}, \downarrow}\right)^2 - \mu \sum_{{\bf r}_i \neq {\bf r}_d}\hat n_{{\bf r}_i,\sigma}~,
\end{aligned}\end{equation}
and ${H}_\text{imp-cbath}$ is the impurity-bath hybridisation:
\begin{equation}\begin{aligned}\label{interactionHamiltonian}
	{H}_\text{imp-cbath} = \frac{J}{\mathcal{Z}}\sum_{\sigma,\sigma^\prime}\sum_{{\bf z}\in\text{NN}({\bf r}_d)} {\bf S}_{{\bf r}_d}\cdot{\boldsymbol \tau}_{\sigma,\sigma^\prime}c^\dagger_{{\bf z},\sigma}c_{{\bf z},\sigma^\prime} \\
    - \frac{V}{\sqrt{\mathcal{Z}}}\sum_{\sigma} \sum_{{\bf z}\in\text{NN}({\bf r}_d)}\left(c^\dagger_{{\bf r}_d,\sigma} c_{{{\bf z}},\sigma} + h.c.\right)~.
\end{aligned}\end{equation}
In this model, ${\bf r}_d$ is the position of the impurity site, and \(\left<{\bf r}_i, {\bf r}_j\right>\neq {\bf r}_d\) indicates that the sum is over all nearest-neighbour pairs of sites avoiding the impurity site \({\bf r}_d\). \(\boldsymbol \tau = \left( \tau_x, \tau_y, \tau_z \right) \) is the vector of Pauli matrices. \(\sigma\) and \(\sigma^\prime\) can be \(\pm 1\) and represent up and down configurations.

The tiled Hamiltonian can be obtained as 
\begin{equation}\begin{aligned}
	\mathcal{H}_\text{tiled} = \mathcal{T}[H({\bf r}_d)]
    % \sum_{{\bf r}}T^\dagger({\bf r} - {\bf r}_d)H({\bf r}_d)T({\bf r} - {\bf r}_d)~,
\end{aligned}\end{equation}
where the $\mathcal{T}[H] = \sum_{{\bf r}}T^\dagger({\bf r} - {\bf r}_d)HT({\bf r} - {\bf r}_d)$ represents the symmetrisation/tiling procedure, and the action of the many-body translation operators $T^\dagger({\bf a})$ on a general operator are defined as
\begin{equation}\begin{aligned}
T^\dagger({\bf a}) \mathcal{O}\left({\bf r}_1, {\bf r}_2,\ldots\right)T({\bf a}) = \mathcal{O}\left({\bf r}_1 + {\bf a}, {\bf r}_2 + {\bf a},\ldots\right)~.
\end{aligned}\end{equation}

We consider the effect of the translation operations on each part of the Hamiltonian:
\begin{equation}\begin{aligned}
	&\mathcal{T}\left[{H}_\text{cbath}({\bf r}_d, {\bf z})\right] = -\frac{(N-2)t}{\sqrt\mathcal{Z}}\sum_{\left<{\bf r}_i, {\bf r}_j\right>;\sigma}\left(c^\dagger_{{\bf r}_i,\sigma}c_{{\bf r}_j,\sigma} + \text{h.c.}\right) \\
    &- \frac{1}{2}W\sum_{\bf r}\left(\hat n_{{\bf r} , \uparrow} - \hat n_{{\bf r} , \downarrow}\right)^2 - \mu (N-1) \sum_{{\bf r}}\hat n_{{\bf r},\sigma}~,\\
	&\mathcal{T}\left[{H}_\text{imp}\right] = -\frac{U}{2}\sum_{{\bf r}}\left(\hat n_{{\bf r} \uparrow} - \hat n_{{\bf r} \downarrow} \right)^2 - \eta\sum_{{\bf r},\sigma} \hat n_{{\bf r}\sigma}~,\\
    &\mathcal{T}\left[{H}_\text{imp-cbath}\right]\\
     &=\sum_{\left< {\bf r}_i, {\bf r}_j\right>}\left[\frac{2}{\mathcal{Z}}J {\bf S}_{{\bf r}_i}\cdot{\bf S}_{{\bf r}_j} - \frac{2}{\sqrt\mathcal{Z}}V \sum_{\sigma}\left(c^\dagger_{{\bf r}_i,\sigma} c_{{{\bf r}_j},\sigma} + h.c.\right)\right]~.
\end{aligned}\end{equation}
We have defined \({\bf S}_{{\bf r}_j} =\sum_{\sigma,\sigma^\prime} {\boldsymbol \tau}_{\sigma,\sigma^\prime}c^\dagger_{{\bf z},\sigma}c_{{\bf z},\sigma^\prime}\) as the local spin operator. While constructing the tiled Hamiltonian, we have added extra copies of the non-interacting Hamiltonian \(\mathcal{H}_\text{cbath-nint} = -\frac{1}{\sqrt\mathcal{Z}}t\sum_{\left<{\bf r}_i, {\bf r}_j\right>;\sigma}\left(c^\dagger_{{\bf r}_i,\sigma}c_{{\bf r}_j,\sigma} + \text{h.c.}\right) - \mu \sum_{{\bf r}}\hat n_{{\bf r},\sigma}\) for the conduction bath (this results in the factors of \(N-2\) and \(N-1\) in front of the first and third terms).

Upon removing these repeated terms, the result of the tiling operations is a Hubbard-Heisenberg model, of the form
\begin{equation}\begin{aligned}
	\mathcal{H}_\text{HH} &= -\frac{1}{\sqrt\mathcal{Z}}\tilde t\sum_{\left<{\bf r}_i, {\bf r}_j\right>;\sigma}\left(c^\dagger_{{\bf r}_i,\sigma}c_{{\bf r}_j,\sigma} + \text{h.c.}\right) - \tilde \mu \sum_{{\bf r}}\hat n_{{\bf r},\sigma} \\
    &+ \frac{1}{\mathcal{Z}}\tilde J\sum_{\left< {\bf r}_i, {\bf r}_j\right>}{\bf S}_{{\bf r}_i}\cdot{\bf S}_{{\bf r}_j} - \frac{1}{2}\tilde U\sum_{\bf r}\left(\hat n_{{\bf r} , \uparrow} - \hat n_{{\bf r} , \downarrow}\right)^2  ~,
\end{aligned}\end{equation}
where the tilde symbol indicates that the parameters are for the lattice model (and not the auxiliary model). By comparing the tiled model and the general lattice model, the lattice model parameters and the auxiliary model parameters can be mapped to each other:
\begin{equation}\begin{aligned}\label{couplingsMappings}
	\tilde t = t+2V,~~ \tilde U = U + W, ~ ~ \tilde \mu = 2\mu + \eta,~ ~ \tilde J = 2J~.
\end{aligned}\end{equation}
In summary, the appropriate method for reconstructing the lattice model Hamiltonian is therefore
\begin{equation}\begin{aligned}\label{tilingPrescriptionFinal}
	\mathcal{H}_\text{tiled} &= \sum_{{\bf r}}\mathcal{H}_\text{aux}({\bf r}) - N\mathcal{H}_\text{cbath-nint}~,
\end{aligned}\end{equation}
where we replaced \(N-3\) with \(N\) assuming a large number of sites. Using this, the extended-SIAM gets ``expanded" into a Hubbard-Heisenberg model.
\end{document}